Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2019-11-16 paper2_JNLS/paper.tex
NM

% Sampling and averaging
\newcommand{\tavg}[1]{\overline{#1}}
\newcommand{\transf}{\mathcal{F}}
\newcommand{\Nsamp}{N_s}
\newcommand{\sumsamp}{\sum_{i=1}^{\Nsamp}}
\newcommand{\Fth}{W}
\newcommand{\RHS}{F}
\newcommand{\RHSh}{\hat{\RHS}}
\newcommand{\uhvec}{\mathbf{u}}
\newcommand{\RHSvec}{\mathbf{F}}
\newcommand{\usqdah}{\hat{v}} %{ (\widehat{u_{da}^2}) }

% Scaling analysis
\newcommand{\traveldist}{\ell}

@@

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC). In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. Inspired by these experimental observations, Majda et al.~\cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In subsequent work, Majda and Qi analyzed more severely truncated systems --- as low as two interacting modes --- that exhibit a statistical phase transition to anomalous statistics and the creation of extreme events while enjoying a more tractable structure \cite{majdaqi2019}. More recently, Qi and Majda demonstrated the capability of machine learning strategies to accurately predict these extreme events \cite{MachineLearning2019}. That work uses a deep neural network coupled with a judicious choice of the entropy loss function that emphasizes the dominant structures of the turbulent field \cite{MachineLearning2019}. It should be noted that, in these studies, extreme events are represented by strong skewness of the wave-field. They occur intermittently and on a relatively frequent basis, which is in contrast to the simpler situation of isolated, rare events \cite{guth2019machine}.

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al.~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. 
In Section \ref{experiments} we detail the laboratory experiments of Bolles et al.~and summarize previously reported findings on anomalous statistics of the surface displacement. In this section, we also present new experimental data on surface {\em slope} statistics --- a line of inquiry motivated by the appearance of the slope in the Hamiltonian structure of TKdV theory developed by Majda et al.~\cite{majda2019}.
In Section \ref{theory}, we detail the TKdV theoretical framework, with a more thorough treatment of the link between model and experimental parameters than was reported in previous work. This link facilitates comparison between the theory and experiments. Within the same TKdV framework, we discuss both a deterministic dynamical systems approach and a statistical mechanics approach. The latter is based on the Hamiltonian structure of TKdV and novel ensemble distributions that incorporate canonical and micro-canonical aspects.
Section \ref{results} presents a systematic comparison between experimental measurements and model predictions, with both surface displacement and surface slope statistics compared closely. This section, also describes deterministic dynamics of TKdV, which offer insight into the experimental length and timescales required for wave statistics to transition. Importantly, this section shows newly collected experimental data related to a a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. The experimental data provides conclusive confirmation of the model prediction. We close with some final remarks in Section \ref{conclusion}.

%confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}. 
% spectacular, remarkable, marvelous, conclusively, convincingly

%We first provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two. Next, we present new experimental data, most notably on {\em surface slope} statistics, which motivates parallel examination of the theory.  Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}.

@@

The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e) on a semi-log scale. The upstream measurements are symmetrically distributed about the mean, $\eta = 0$. In fact, Bolles et al.~found that these measurements follow a Gaussian distribution closely \cite{bolles2019}. The downstream measurements, however, skew strongly towards positive displacement, $\eta > 0$. Bolles et al.~found these measurements to be well described by a mean-zero gamma distribution \cite{bolles2019}. The slower decay rate of this downstream distribution indicates an elevated level of extreme surface displacement, i.e.~rogue waves. Bolles et al.~estimated that a rogue wave can be up to 65 times more likely in these experiments than if displacements were Gaussian \cite{bolles2019}.

@@

	Figure \ref{ExpSpatialStats} shows how these statistics vary in the vicinity of the ADC, located at $x = 0$, for both displacement and slope. First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves. Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC. The skewness and kurtosis, however, respond strongly to the ADC as long as the amplitude is sufficiently high. As seen in \ref{ExpSpatialStats}(b)--(c), both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm. Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude. The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e). Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution. We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.

@@

	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}. Our attention to slope statistics was motivated by new theoretical developments by Majda et al.~\cite{majda2019} and will be expanded upon in later sections. We extract the surface slope by numerical differentiating images of the free surface using Savitzky-Golay smoothing filters. This ability to extract surface slope is another advantage of our optical measurements over the more commonly used technique of placing a set of discrete wave probes in the tank. As seen in Fig.~\ref{ExpSpatialStats}(a) the standard deviation of slope behaves similar to displacement: $\std(\eta_x)$ increases with driving amplitude, but remains nearly uniform in space for each individual experiment and is not affected by the ADC. Once again though, the higher-order moments respond strongly to the ADC if the driving amplitude is sufficiently high. The skewness of slope becomes highly negative and reaches a minimum near $x = 8$ cm, which is downstream of the ADC but upstream of the position where the displacement statistics peak. The kurtosis of $\eta_x$ reaches a peak at nearly the same location as skewness. We note that the large {\em negative} skewness of $\eta_x$ indicates a bias towards negative slope, which is consistent with a right-moving wave of steep leading surface and shallower trailing surface, i.e.~a wave that is near overturning.

	Bolles et al.~systematically explored the effect of varying the driving amplitude on the anomalous statistics for a fixed depth ratio. 
Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we also report experiments in which the depth ratio was varied. 
To have a simple baseline for comparison against theory, we refer to the data from Fig.~\ref{ExpDiagStats} as the {\em reference} experiments. In these experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the two largest amplitudes shown in Fig.~\ref{ExpSpatialStats}), the upstream depth is $\dup = 12.5$ cm, and the downstream depth is $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$. From the data represented in Fig.~\ref{ExpSpatialStats}, we extract a characteristic wave amplitude for the representative experiments of $\etastd = 0.21$ cm, which will be important for setting dimensionless parameters that enter the theory. Table \ref{paramtable} lists the range of experimental parameters along with these representative values, as well as values of dimensionless parameters that will be introduced later.
% NOTE: Maybe modify slightly the last paragraph if it does not seem to fit next time I read it.

@@

In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from Majda et al.~2019 \cite{majda2019}. Since the choice of normalization is not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices. To this end, we consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment. We introduce the dimensionless variables
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = (x-ct) / \lengthscale
&&\mbox{\em dimensionless position (in moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}
We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.

@@

We remark that, in defining the Hamiltonian, we have chosen the sign convention of Lax \cite{lax1975periodic}. More recent work of Bajars et al.~\cite{bajars2013weakly} and Majda et al.~\cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite. Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent. Using the second convention, Majda et al.~found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}. We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.

@@

	In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced by Abramov et al.~for the Burgers-Hopf system \cite{abramov2003}. Specifically, this ensemble is microcanonical in energy and canonical in the Hamiltonian. The reason this ensemble is needed is the sign indefiniteness of the cubic term $\Hthree$ in the Hamiltonian, which would cause a simple canonical distribution to diverge at infinity. The mixed ensemble, however, fixes the energy and hence confines dynamics to the compact set of the unit hypersphere $\En = 1$. Since the Hamiltonian is continuous, its value is bounded on the unit hypersphere, and thus the mixed ensemble produces a normalizable distribution. This construction applies equally well to the truncated or untruncated KdV system and hence we will not distinguish between the two. We note that other possible constructions may be applicable too \cite{kleeman2014nonequilibrium}.

@@

	This statistical matching condition, originally derived by Majda et al.~(2019) \cite{majda2019}, imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$. In particular, we view $\thup$ as given by the random-state of the incoming wave field. Thus, while we will treat $\thup$ as a free parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by the matching condition \eqref{statmatch}, giving the functional dependence
\begin{equation}
\thdn = \transf \left( \thup \right)
\qquad \mbox{\em Transfer function}
\end{equation}
The {\em transfer function} $\transf$ will be a key link needed to relate the theory to experiments.

@@

We now discuss arguably the most novel single result of the manuscript. In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}. More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$. Here, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.

@@

	This manuscript extends the parallel experimental and modeling efforts of Bolles et al.~\cite{bolles2019} and Majda et al.~\cite{majda2019} concerning the emergence of anomalous wave statistics from abrupt changes in bottom topography. The theoretical framework is based on deterministic and statistical analysis of the TKdV system, with exploitation of the Hamiltonian structure and the associated invariant Gibbs measures. The theory depends crucially on matching the incoming and outgoing invariant measures at the abrupt depth change, so as to link the incoming and outgoing states. Throughout, we have emphasized the synergy between the experiments and theory, with experimental data informing theoretical advancements and model predictions motivating new experimental measurements.

@@

	Finally, Majda et al.~\cite{majda2019} derived an explicit formula for the skewness of the outgoing wave field, which we have tested against direct experimental measurements. Specifically, the formula predicts how the ratio of displacement skewness to slope variance, $\skw(\eta) / \Delta \var(\eta_x)$, depends on wave amplitude. The experimental measurements conclusively confirm the inverse-cube dependence on wave amplitude, thus highlighting the predictive power of the TKdV statistical mechanics framework.

>>>>

% Sampling and averaging
\newcommand{\tavg}[1]{\overline{#1}}
\newcommand{\transf}{\mathcal{F}}
\newcommand{\Nsamp}{N_s}
\newcommand{\sumsamp}{\sum_{i=1}^{\Nsamp}}
\newcommand{\Fth}{W}
\newcommand{\RHS}{F}
\newcommand{\RHSh}{\hat{\RHS}}
\newcommand{\uhvec}{\mathbf{u}}
\newcommand{\RHSvec}{\mathbf{F}}
\newcommand{\usqdah}{\hat{v}} %{ (\widehat{u_{da}^2}) }
% Scaling analysis
\newcommand{\traveldist}{\ell}
% et al
\newcommand{\boetal}{Bolles {\it et al.}~(2019)~}
\newcommand{\maetal}{Majda {\it et al.}~(2019)~}

@@

Our particular focus is the laboratory experiments of \boetal~\cite{bolles2019}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC). In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. Inspired by these experimental observations, \maetal~\cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In subsequent work, Majda and Qi analyzed more severely truncated systems --- as low as two interacting modes --- that exhibit a statistical phase transition to anomalous statistics and the creation of extreme events while enjoying a more tractable structure \cite{majdaqi2019}. More recently, Qi and Majda demonstrated the capability of machine learning strategies to accurately predict these extreme events \cite{MachineLearning2019}. That work uses a deep neural network coupled with a judicious choice of the entropy loss function that emphasizes the dominant structures of the turbulent field \cite{MachineLearning2019}. It should be noted that, in these studies, extreme events are represented by strong skewness of the wave-field. They occur intermittently and on a relatively frequent basis, which is in contrast to the simpler situation of isolated, rare events \cite{guth2019machine}.

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of \boetal~\cite{bolles2019} and \maetal~\cite{majda2019, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. 
In Section \ref{experiments} we detail the laboratory experiments of \boetal~and summarize previously reported findings on anomalous statistics of the surface displacement. In this section, we also present new experimental data on surface {\em slope} statistics --- a line of inquiry motivated by the appearance of the slope in the Hamiltonian structure of TKdV theory developed by \maetal~\cite{majda2019}.
In Section \ref{theory}, we detail the TKdV theoretical framework, with a more thorough treatment of the link between model and experimental parameters than was reported in previous work. This link facilitates comparison between the theory and experiments. Within the same TKdV framework, we discuss both a deterministic dynamical systems approach and a statistical mechanics approach. The latter is based on the Hamiltonian structure of TKdV and novel ensemble distributions that incorporate canonical and micro-canonical aspects.
Section \ref{results} presents a systematic comparison between experimental measurements and model predictions, with both surface displacement and surface slope statistics compared closely. This section, also describes deterministic dynamics of TKdV, which offer insight into the experimental length and timescales required for wave statistics to transition. Importantly, this section shows newly collected experimental data related to a a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. The experimental data provides conclusive confirmation of the model prediction. We close with some final remarks in Section \ref{conclusion}.

%confirms the predictive power of the framework developed by \maetal~\cite{majda2019}. 
% spectacular, remarkable, marvelous, conclusively, convincingly

%We first provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two. Next, we present new experimental data, most notably on {\em surface slope} statistics, which motivates parallel examination of the theory.  Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. Comparison with the new measurements further confirms the predictive power of the framework developed by \maetal~\cite{majda2019}.

@@

The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e) on a semi-log scale. The upstream measurements are symmetrically distributed about the mean, $\eta = 0$. In fact, \boetal found that these measurements follow a Gaussian distribution closely \cite{bolles2019}. The downstream measurements, however, skew strongly towards positive displacement, $\eta > 0$. \boetal found these measurements to be well described by a mean-zero gamma distribution \cite{bolles2019}. The slower decay rate of this downstream distribution indicates an elevated level of extreme surface displacement, i.e.~rogue waves. \boetal estimated that a rogue wave can be up to 65 times more likely in these experiments than if displacements were Gaussian \cite{bolles2019}.

@@

	Figure \ref{ExpSpatialStats} shows how these statistics vary in the vicinity of the ADC, located at $x = 0$, for both displacement and slope. First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves. Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC. The skewness and kurtosis, however, respond strongly to the ADC as long as the amplitude is sufficiently high. As seen in \ref{ExpSpatialStats}(b)--(c), both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm. Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude. The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e). \boetal found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution. We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.

@@

	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}. Our attention to slope statistics was motivated by new theoretical developments by \maetal \cite{majda2019} and will be expanded upon in later sections. We extract the surface slope by numerical differentiating images of the free surface using Savitzky-Golay smoothing filters. This ability to extract surface slope is another advantage of our optical measurements over the more commonly used technique of placing a set of discrete wave probes in the tank. As seen in Fig.~\ref{ExpSpatialStats}(a) the standard deviation of slope behaves similar to displacement: $\std(\eta_x)$ increases with driving amplitude, but remains nearly uniform in space for each individual experiment and is not affected by the ADC. Once again though, the higher-order moments respond strongly to the ADC if the driving amplitude is sufficiently high. The skewness of slope becomes highly negative and reaches a minimum near $x = 8$ cm, which is downstream of the ADC but upstream of the position where the displacement statistics peak. The kurtosis of $\eta_x$ reaches a peak at nearly the same location as skewness. We note that the large {\em negative} skewness of $\eta_x$ indicates a bias towards negative slope, which is consistent with a right-moving wave of steep leading surface and shallower trailing surface, i.e.~a wave that is near overturning.

	\boetal systematically explored the effect of varying the driving amplitude on the anomalous statistics for a fixed depth ratio. 
Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we also report experiments in which the depth ratio was varied. 
To have a simple baseline for comparison against theory, we refer to the data from Fig.~\ref{ExpDiagStats} as the {\em reference} experiments. In these experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the two largest amplitudes shown in Fig.~\ref{ExpSpatialStats}), the upstream depth is $\dup = 12.5$ cm, and the downstream depth is $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$. From the data represented in Fig.~\ref{ExpSpatialStats}, we extract a characteristic wave amplitude for the representative experiments of $\etastd = 0.21$ cm, which will be important for setting dimensionless parameters that enter the theory. Table \ref{paramtable} lists the range of experimental parameters along with these representative values, as well as values of dimensionless parameters that will be introduced later.
% NOTE: Maybe modify slightly the last paragraph if it does not seem to fit next time I read it.

@@

In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from \maetal \cite{majda2019}. Since the choice of normalization is not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices. To this end, we consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment. We introduce the dimensionless variables
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = (x-ct) / \lengthscale
&&\mbox{\em dimensionless position (in moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}
We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.

@@

We remark that, in defining the Hamiltonian, we have chosen the sign convention of Lax (1975) \cite{lax1975periodic}. More recent work of Bajars {\it et al.}~(2013) \cite{bajars2013weakly} and \maetal \cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite. Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent. Using the second convention, \maetal found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}. We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.

@@

	In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced by Abramov {\it et al.}~(2003) for the Burgers-Hopf system \cite{abramov2003}. Specifically, this ensemble is microcanonical in energy and canonical in the Hamiltonian. The reason this ensemble is needed is the sign indefiniteness of the cubic term $\Hthree$ in the Hamiltonian, which would cause a simple canonical distribution to diverge at infinity. The mixed ensemble, however, fixes the energy and hence confines dynamics to the compact set of the unit hypersphere $\En = 1$. Since the Hamiltonian is continuous, its value is bounded on the unit hypersphere, and thus the mixed ensemble produces a normalizable distribution. This construction applies equally well to the truncated or untruncated KdV system and hence we will not distinguish between the two. We note that other possible constructions may be applicable too \cite{kleeman2014nonequilibrium}.

@@

	This statistical matching condition, originally derived by \maetal \cite{majda2019}, imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$. In particular, we view $\thup$ as given by the random-state of the incoming wave field. Thus, while we will treat $\thup$ as a free parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by the matching condition \eqref{statmatch}, giving the functional dependence
\begin{equation}
\thdn = \transf \left( \thup \right)
\qquad \mbox{\em Transfer function}
\end{equation}
The {\em transfer function} $\transf$ will be a key link needed to relate the theory to experiments.

@@

We now discuss arguably the most novel single result of the manuscript. In the recent theoretical study, \maetal derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}. More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$. Here, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.

@@

	This manuscript extends the parallel experimental and modeling efforts of \boetal \cite{bolles2019} and \maetal \cite{majda2019} concerning the emergence of anomalous wave statistics from abrupt changes in bottom topography. The theoretical framework is based on deterministic and statistical analysis of the TKdV system, with exploitation of the Hamiltonian structure and the associated invariant Gibbs measures. The theory depends crucially on matching the incoming and outgoing invariant measures at the abrupt depth change, so as to link the incoming and outgoing states. Throughout, we have emphasized the synergy between the experiments and theory, with experimental data informing theoretical advancements and model predictions motivating new experimental measurements.

@@

	Finally, \maetal \cite{majda2019} derived an explicit formula for the skewness of the outgoing wave field, which we have tested against direct experimental measurements. Specifically, the formula predicts how the ratio of displacement skewness to slope variance, $\skw(\eta) / \Delta \var(\eta_x)$, depends on wave amplitude. The experimental measurements conclusively confirm the inverse-cube dependence on wave amplitude, thus highlighting the predictive power of the TKdV statistical mechanics framework.
